---
title: "Maryland Bridges"
author: "Giedrius Blazys"
date: "11/30/2018"
description: 'A list of resources for working with databases and Rxx.'
slug: baltimore-bridges
categories:
  - R
  - TidyTuesday
tags:
  - '#rstats'
  - '#tidytuesday'
output:
  html_document:
    code_folding: hide
    fig_height: 4
    fig_width: 7
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 120)
```

# {.tabset}

## Description

[youtube link](https://www.youtube.com/watch?v=kzM-4jMh9Qs)

[#tidytuesday project repository](https://github.com/rfordatascience/tidytuesday)

Lets first have a look at these files:

## Load packages and data

```{r load_libraries, message=FALSE, warning=FALSE}
library("tidyverse")
#library("ggthemes")
library("scales")
theme_set(theme_light())
```


```{r load_data, message=FALSE, warning=FALSE}
data_file_link <- "https://github.com/rfordatascience/tidytuesday/raw/master/data/2018-11-27/baltimore_bridges.csv"
maryland_bridges <- read_csv(data_file_link) %>% 
    select(-vehicles) %>% 
    mutate(inspection_yr = inspection_yr + 2000,
           decade = 10 * (yr_built %/% 10)) %>% 
    filter(yr_built >= 1900) %>% 
    replace_na(list(responsibility = "Other")) %>% 
    mutate(responsibility = fct_lump(responsibility, 4),
           county = str_to_title(county))
```

## Analysis

When briges were build?

```{r}
maryland_bridges  %>% 
    count(decade, sort = TRUE) %>% 
    ggplot(aes(decade, n))+geom_line() + 
    expand_limits(y=0)+
    labs(y = "# of bridges built this decade")
```

How bridge conditions depend on the year built?

```{r}
maryland_bridges  %>% 
    group_by(decade) %>% 
    summarise(pct_good = mean(bridge_condition == "Good"),
              total = n()) %>% 
    ggplot(aes(decade, pct_good))+geom_line() + 
    geom_smooth(method = "lm")+
    scale_y_continuous(labels = percent_format())+
    expand_limits(y=0)+
    labs(y = "% of bridges in good condition")
```

Most bridges build before 1970 not in good condition (though it does not matter when before 1970), but most of bridges build after 1980 are in good condition.

```{r}
maryland_bridges  %>% 
    count(responsibility, sort = TRUE) %>% 
    mutate(responsibility = fct_reorder(responsibility, n)) %>% 
    ggplot(aes(responsibility, n)) + 
    geom_col() + 
    coord_flip()

```

```{r}
maryland_bridges  %>% 
    group_by(responsibility, decade) %>% 
    summarise(pct_good = mean(bridge_condition == "Good"),
              total = n()) %>% 
    filter(responsibility != "Other") %>% 
    ggplot(aes(decade, pct_good, color = responsibility))+geom_line() + 
    scale_y_continuous(labels = percent_format())+
    expand_limits(y=0)+
    labs(y = "% of bridges in good condition")
```

The Country Highway Agency consistently has the most bridges rated as "Good" from before 1970. State Toll Authority is underperforming in bridges built in the 1990s.

```{r}
maryland_bridges  %>% 
    ggplot(aes(avg_daily_traffic)) +
    geom_histogram(bins = 30) +
    scale_x_log10(labels = comma_format())
```

```{r}
maryland_bridges  %>% 
    filter(yr_built >= 1990) %>% 
    group_by(traffic_category = cut(avg_daily_traffic, c(0,1000,10000,Inf),
             labels = c("<1,000", "1,000-10,000", "10,000+"))) %>% 
    summarise(pct_good = mean(bridge_condition == "Good"),
              total = n())
    
```

```{r}
maryland_bridges  %>% 
    ggplot(aes(long, lat, color = responsibility)) + 
    borders("state", regions = "Maryland") +
    geom_point() + 
    coord_map() + 
    theme_void()
```

What does the traffic look geographically

```{r}
maryland_bridges  %>% 
    ggplot(aes(long, lat, color = avg_daily_traffic)) + 
    borders("state", regions = "Maryland") +
    geom_point() + 
    scale_color_gradient2(low = "blue", high = "red", 
                          midpoint = log10(median(maryland_bridges$avg_daily_traffic)),
                          trans = "log10",
                          labels = comma_format()) +
    coord_map() + 
    theme_void()
```

```{r}
maryland_bridges  %>% 
    ggplot(aes(long, lat, color = bridge_condition)) + 
    borders("state", regions = "Maryland") +
    geom_point(size = 1) + 
    coord_map() + 
    theme_void()
```

```{r}
maryland_bridges  %>% 
    ggplot(aes(long, lat, color = decade)) + 
    borders("state", regions = "Maryland") +
    geom_point(size = 1) + 
    coord_map() + 
    theme_void()
```


```{r}
maryland_bridges  %>% 
    ggplot(aes(long, lat, color = county)) + 
    borders("state", regions = "Maryland") +
    geom_point(size = 1) + 
    coord_map() + 
    theme_void()
```

### Effect of country, responsibility, traffic and time on bridge condition

```{r}
library(broom)
library(splines)
```

Controlled for the overall trend in change over time, what's the effect of being in a particular responsibility

```{r}
model1 <- maryland_bridges %>% 
    mutate(good = bridge_condition == "Good") %>% 
    glm(good ~ ns(yr_built, 4) + responsibility, data = ., family = "binomial") 

augment(model1, maryland_bridges, type.predict = "response") %>% 
    ggplot(aes(yr_built, .fitted, color = responsibility)) + 
    geom_line() +
    expand_limits(y=0) +
    scale_y_continuous(labels = percent_format()) +
    labs(y = "Predicted probability a bridge is rated 'good'")
```

```{r}
model1 %>% 
    tidy(conf.int = TRUE) %>% 
    filter(str_detect(term, "responsibility")) %>% 
    mutate(term = reorder(term, estimate)) %>% 
    ggplot(aes(estimate, term)) + 
    geom_point() + 
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
    geom_vline(color = "red", xintercept = 0, lty = 2)
```

Controlled for the overall trend in change over time, what's the effect of being in a particular country or responsibility

```{r}
model2 <- maryland_bridges %>% 
    mutate(good = bridge_condition == "Good") %>% 
    glm(good ~ ns(yr_built, 4) + responsibility + county, data = ., family = "binomial") 

augment(model2, maryland_bridges, type.predict = "response") %>% 
    ggplot(aes(yr_built, .fitted, color = responsibility)) + 
    geom_line() +
    facet_wrap(~ county) +
    expand_limits(y=0) +
    scale_y_continuous(labels = percent_format()) +
    labs(y = "Predicted probability a bridge is rated 'good'")
```

```{r}
model2 %>% 
    tidy(conf.int = TRUE) %>% 
    filter(str_detect(term, "responsibility|county")) %>% 
    mutate(term = reorder(term, estimate)) %>% 
    ggplot(aes(estimate, term)) + 
    geom_point() + 
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
    geom_vline(color = "red", xintercept = 0, lty = 2)
```

```{r}
model3 <- maryland_bridges %>% 
    mutate(good = bridge_condition == "Good") %>% 
    glm(good ~ ns(yr_built, 4) + responsibility + ns(avg_daily_traffic,3), data = ., family = "binomial") 

model3 %>% summary()
```

```{r}
model2 %>% summary()
```

